Multi-domain, multilingual translation with latent multi-task group dropout

### 2021-11-15 2022/acl/acl_latex.tex
answer fyTodos, correct spelling errors in abstract

\begin{abstract}
  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalties of the tasks at hand, while smaller parts are reserved to model the peculariaties of a language or a domain. In adapter-based approaches, these strategies are hard coded in the network architecture, independent of the similarities that exist between tasks. In this work, we propose a new method aimed at better taking advantage of these similarities, using a latent-variable model. We also develop new techniques to train this model end to end and report multiple experimental results that show the the learned patterns are both meaningful and yield improved translation performance, without any increase of the model size. 
\end{abstract}

@@

\subsection{System design and configuration}
\subsubsection{Multidomain translation systems}
Our systems for the multidomain experiments are designed as follows:
\begin{itemize}
\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads;
\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
\item \system{LaMGD} Transformer:\fyTodo{Explain acronym} There is no change in the architecture. We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 out of the 16 groups are selected. The number of parameters of the variational distributions is $k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
\end{itemize}

@@

All the translation systems are implemented OpenNMT\fyTodo{Fix this, add reference to framework if needed}.

@@

For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \cite{Post18call}.\fyTodo{Check this}

\subsection{Datasets}
\subsubsection{Multidomain translation}
We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyTodo{Add public to the data link} The datasets for the multidomain translation experiments are detailled in Table~\ref{tab:Corpora-chap4}.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
\begin{itemize}
\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
\end{itemize}

The languages used in multilingual experiments are refered to by their ISO 639-1 code that we reproduce below for convenience:
\begin{itemize}
\item \textbf{Diverse set}: bos (Bosnian), Bulgarian (bul), French (fra), ell (Greek),
  hin (Hindi), Korean (kor) mkd (Macedonian), mar (Marathi);
\item \textbf{Related set}: Azerbajiani (aze), Belarusian (bel),
  Czech (ces), Galician (glg), Portuguese (por), Russian (rus), Slovak (slk), Turkish (tur).
\end{itemize}

Table~\ref{tab:Corpora-multilingual} describes the subsets of TED8 that were used in our experiments.
\begin{table*}[h!]
  \centering
  \begin{tabular}{l|cccc} 
  \hline
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
    \hline 
    \multirow{8}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 \\
							 & Belarusian & 4.51k & 248 & 664 \\
							 & Glacian & 10.0k & 682 & 1007 \\
							 & Slovak & 61.5k & 2271 & 2445 \\
							 & Turkish & 182k & 4045 & 5029 \\
                             & Russian & 208k & 4814 & 5483 \\
                             & Portuguese & 185k & 4035 & 4855 \\
                             & Czech & 103k & 3462 & 3831 \\
    \hline
    \multirow{8}{*}{Related} & Bosnian & 5.64k & 474 & 463 \\
                             & Marathi & 9.84k & 767 & 1090 \\
                             & Hindi & 18.79k & 854 & 1243 \\
                             & Macedonian & 25.33k & 640 & 438 \\
                             & Greek & 134k & 3344 & 4433 \\
                             & Bulgarian & 174k & 4082 & 5060 \\
                             & French & 192k & 4320 & 4866 \\
                             & Korean & 205k & 4441 & 5637 \\
  \end{tabular}
  \caption{Data Statistics of TED8 Datasets}
\label{tab:Corpora-multilingual}
\end{table*}

\section{Results and analyses}

@@

\subsection{Multilingual translation}
Results for the multilingual experiments are in Table~\ref{tab:multilingual}.  The \system{LaMGD} Tranformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively. Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets. However, \system{LaMGD} Tranformer is outperformed by the multilingual \system{Transformer} and Language Adapters for the O2M-diverse condition.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Experiments with multilingual translation.\fyTodo{What is the bold for ?}}
  \label{tab:multilingual}
\end{table*}

@@

We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by PCA.
\begin{figure*}[htbp]
\begin{subfigure}{0.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.85\textwidth]{ted_related}  
  \caption{TED-Related}
  \label{fig:visual-related}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.85\textwidth]{ted_diverse}  
  \caption{TED-Diverse}
  \label{fig:visual-diverse}
\end{subfigure}
\caption{Visualization of languages according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:visual}
\end{figure*}

@@

Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer. This method is closely related to the FusionAdapters of \citet{pfeiffer21adapterfusion}. Both approaches assume that the final representation\fyTodo{??} is a combination or a choice between shared representation and task-specific representations. The proximity between tasks is not taken into account in the routing mechanism. Our method proposed a different approach to the problem of multi-task routing in the underlying network.

\section{Conclusions and outlook}\label{sec:conclusion}
In this work, we have presented a novel method for multdomain and multilingual translation. It allows us to jointly search for an optimal assignment of sub-networks to tasks and to learn the parameters of the underlying network. Our method relies on a sound mathematical framework and an end-to-end optimization procedure which implies very little extra parameters.\fyTodo{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters. For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 cases. Besides, we provided an thorough analysis of the similarities between learned sub-networks and demonstrated a strong correlation between the sub-networks' similarity and the proximity of the corresponding tasks (domain or language).

@@

\begin{align*}
&\frac{\partial g}{\partial \nu} \times \frac{\partial \nu}{\partial \Phi_l^d} + \frac{\partial g}{\partial \Phi_l^d} = 0 \\
& \Rightarrow \frac{\partial \nu}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \nu}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}
\end{align*}

@@

\section{Appendix B}
\label{appendix:b}
In this section, we give a simple proof of Inequality \ref{eq:entropy}. In fact, we only need to prove $\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] \geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]$. The proof is as follows
\begin{align*}
&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
&\geqslant -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1 | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
\end{align*}
\section{Appendix C}
\begin{figure*}
\includegraphics[width=0.7\textwidth]{multi_domain}
\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:domain}
\end{figure*}

\section{Appendix D}
\begin{breakablealgorithm}
\caption{Training \system{LaMGD}} \label{alg:mdmt}
\begin{algorithmic}[1]
\REQUIRE {
  \begin{itemize}
    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
  \item[] 
  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
  \item number of retained groups: k, number of groups: $n_p$
 \end{itemize}}
\REPEAT 
\STATE{Pick a batch from domain $d$}
\STATE{Sampling $\forall p,  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
\IF{$Early\_stopping(Eval\_scores)$}
	\STATE{break}
\ENDIF
\STATE{$i=i+1$}
\UNTIL{$i> iter\_num$}
\end{algorithmic}
\end{breakablealgorithm}

\end{document}

>>>>

\begin{abstract}
  Multidomain and multilingual machine translation often rely on parameter sharing strategies, where large portions of the network are meant to capture the commonalities of the tasks at hand, while smaller parts are reserved to model the peculiarities of a language or a domain. In adapter-based approaches, these strategies are hardcoded in the network architecture, independent of the similarities between tasks. In this work, we propose a new method to better take advantage of these similarities, using a latent-variable model. We also develop new techniques to train this model end to end and report multiple experimental results that show the learned patterns are both meaningful and yield improved translation performance without any increase of the model size. 
\end{abstract}

@@

\subsection{System design and configuration}
\subsubsection{Multidomain translation systems}
Our systems for the multidomain experiments are designed as follows:
\begin{itemize}
\item Transformer: The embedding dimension for both encoder and decoder is set as 512, and the feedforward dimension is 2048; the multi-head attention mechanism contains 8 heads;
\item Adapter-based Transformer: The intermediate feedforward dimension is set to 2048;% as in \citet{Pham21revisiting}.
\item Transformer using Latent multi-task group dropout (denoted \system{LaMGD} Transformer):\fyDone{Explain acronym} There is no change in the architecture. We group the 512 nodes in each layer into 16 groups of 32 consecutive nodes. For each domain, only 12 out of the 16 groups are selected. The number of parameters of the variational distributions is $k \times L \times n_d$, which is negligible in comparison to the size of the Transformer model.
\end{itemize}

@@

All the translation systems are implemented with OpenNMT-tf \footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}\fyDone{Fix this, add reference to framework if needed}.

@@

For all experiments, we report the BLEU score of \citet{Papineni02bleu} computed with SacreBleu \cite{Post18call}.\fyDone{Check this}

\subsection{Datasets}
\subsubsection{Multidomain translation}
We use the same data as in the recent work of \citet{Pham21revisiting} on multidomain translation.\fyDone{Add public to the data link} The datasets \footnote{The raw data can be downloaded from \url{https://github.com/qmpham/experiments/tree/main/tacl20}} for the multidomain translation experiments are detailed in Table~\ref{tab:Corpora-chap4}.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|l|cccccc|} %*{4}{|r|}}
    \cline{2-7} 
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) \\
    \# \revisiondone{tokens}  &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 \\
    \# \revisiondone{types}  & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 \\
    \# \revisiondone{uniq} & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the basic domain mixture (which does not include the \domain{news} domain), number of tokens in English and French ($\times 10^6$), number of types in English and French ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora-chap4}
\end{table*}
\subsubsection{Multilingual translation}
We evaluate our model on both one-to-many (O2M) and many-to-one (M2O)
translation tasks borrowing the multilingual translation datasets from past studies. More precisely, we used: 
\begin{itemize}
\item TED8-Related. Following the setting of \citet{Wang20balancing}, we use a subset of translations from \citet{qi18when} between English and eight \emph{related languages}.
\item TED8-Diverse. The dataset consists of parallel sentences between English and eight \emph{diverse languages} as in \citet{Wang20balancing}.
\end{itemize}

The languages used in multilingual experiments are referred to by their ISO 639-1 code that we reproduce below for convenience:
\begin{itemize}
\item \textbf{Diverse set}: bos (Bosnian), Bulgarian (bul), French (fra), ell (Greek),
  hin (Hindi), Korean (kor) mkd (Macedonian), mar (Marathi);
\item \textbf{Related set}: Azerbajiani (aze), Belarusian (bel),
  Czech (ces), Galician (glg), Portuguese (por), Russian (rus), Slovak (slk), Turkish (tur).
\end{itemize}

Table~\ref{tab:Corpora-multilingual} describes the subsets of TED8 that were used in our experiments.
\begin{table*}[h!]
  \centering
  \begin{tabular}{l|cccc|l|cccc} 
  \hline
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} & \multicolumn{1}{|c|}{} & \multicolumn{1}{c}{\domain{lang}} & \multicolumn{1}{c}{\domain{train}} & \multicolumn{1}{c}{\domain{dev}} & \multicolumn{1}{c}{\domain{test}} \\
    \hline 
    \multirow{2}{*}{Related} & Azerbaijani & 5.94k & 671 & 903 & \multirow{2}{*}{Diverse} & Bosnian & 5.64k & 474 & 463 \\
							 & Belarusian & 4.51k & 248 & 664 & & Marathi & 9.84k & 767 & 1090\\
							 & Glacian & 10.0k & 682 & 1007 & & Hindi & 18.79k & 854 & 1243\\
							 & Slovak & 61.5k & 2271 & 2445 & & Macedonian & 25.33k & 640 & 438 \\
							 & Turkish & 182k & 4045 & 5029 & & Greek & 134k & 3344 & 4433\\
                             & Russian & 208k & 4814 & 5483 & & Bulgarian & 174k & 4082 & 5060 \\
                             & Portuguese & 185k & 4035 & 4855 & & French & 192k & 4320 & 4866\\
                             & Czech & 103k & 3462 & 3831 & & Korean & 205k & 4441 & 5637 \\
    \hline
  \end{tabular}
  \caption{Data Statistics of TED8 Datasets}
\label{tab:Corpora-multilingual}
\end{table*}
\section{Results and analyses}

@@

\subsection{Multilingual translation}
Results for the multilingual experiments are in Table~\ref{tab:multilingual}.  The \system{LaMGD} Tranformer achieves an improvement of 0.42, 0.33, 0.32 in average over the multilingual \system{Transformer} in the O2M-related, M2O-related, M2O-diverse conditions, respectively. Significant gains are observed for languages \domain{bel}, \domain{glg} (both direction), \domain{hin} and \domain{bos} (O2M direction) which are very low-resource languages in our sets. However, \system{LaMGD} Tranformer is outperformed by the multilingual \system{Transformer} and Language Adapters for the O2M-diverse condition.
\begin{table*}[h!]
  \centering
  \begin{tabular}{|p{4cm}|*{9}{r|}} \hline
    O2M-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[91.6m]} & 4.8 &7.3&20.8&21.1&39.7&19.8&22.6&15.2&18.9 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &4.3&6.8&21.1&22&39.7&20&23&15.2&19 \\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]}  & 5.2&\SB{9.4}&20.6&\SB{22.8}&39.6&19.6&22.4&15.0&19.33 \\ 
	\hline
    \hline
    M2O-related & \multicolumn{1}{c|}{\domain{aze}} & \multicolumn{1}{c|}{\domain{ bel}} & \multicolumn{1}{c|}{\domain{ces}} & \multicolumn{1}{c|}{\domain{glg}} & \multicolumn{1}{c|}{\domain{por}} & \multicolumn{1}{c|}{\domain{rus}} & \multicolumn{1}{c|}{\domain{slk}} & \multicolumn{1}{c|}{\domain{tur}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[67.8m]} &11.4&16.6&28.5&	27.1&43.7&24.6&30.3&25.6&25.98 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]} &10.1&15.8&28.4&26.8&43.7&24.5&30.2&25.6&25.64\\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}  &11.3&\SB{17.4}&28.6&\SB{28.7}&43.7&24.5&30.7&25.6&26.31 \\ 
    \hline
    \hline
    O2M-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[96.9m]} & 10.2&4&12.7&22.2&31.8&\SB{34.0}&38.3&8.3&20.19 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &10.2&4&\SB{13.3}&21.9&32.2&\SB{34.1}&38.5&8.3&20.31 \\ 
    \system{LaMGD} Tranformer   \hfill{\footnotesize[+0m]}&10.1&3.8&12.6&\SB{22.8}&31.8&33.4&38.1&8.1&20.09\\
    \hline 
    \hline
    M2O-diverse & \multicolumn{1}{c|}{\domain{bos}} & \multicolumn{1}{c|}{\domain{mar}} & \multicolumn{1}{c|}{\domain{hin}} & \multicolumn{1}{c|}{\domain{mkd}} & \multicolumn{1}{c|}{\domain{ell}} & \multicolumn{1}{c|}{\domain{bul}} & \multicolumn{1}{c|}{\domain{fra}} & \multicolumn{1}{c|}{\domain{kor}} & \multicolumn{1}{c|}{\domain{avg}} \\ \hline 
    \system{Transformer}  \hfill{\footnotesize[70.4m]} &22.4&9.7&20.5&31.8&37.5&38.7&39.8&19.0&27.43 \\
    \revision{\system{Adapter}}   \hfill{\footnotesize[13m]}  &22.5&9.4&20.0&30.6&37.2&38.2&39.3&19.0&27.03\\ 
    \system{LaMGD} Tranformer  \hfill{\footnotesize[+0m]} &\SB{23.5}&9.6&\SB{21.5}&32.2&37.7&38.6&40.0&18.9&27.75 \\
    \hline
  \end{tabular}
  \caption{Experiments with multilingual translation. Boldface denotes significant gains over \system{Transformer}\fyDone{What is the bold for ?}}
  \label{tab:multilingual}
\end{table*}

@@

We also plot the languages based on their dropout masks in Figure~\ref{fig:visual} using a 2d projection computed by PCA.
\begin{figure*}[htbp]
\begin{subfigure}{0.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=0.9\textwidth]{ted_related}  
  \caption{TED-Related}
  \label{fig:visual-related}
\end{subfigure}
\begin{subfigure}{0.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=0.85\textwidth]{ted_diverse}  
  \caption{TED-Diverse}
  \label{fig:visual-diverse}
\end{subfigure}
\caption{Visualization of languages according to their dropout masks (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:visual}
\end{figure*}

@@

Recently, \citet{biao21share} proposed jointly learning how to route between shared representation and language-specific representation via a conditional language-specific routing and the parameters of the underlying Transformer. This method is closely related to the FusionAdapters of \citet{pfeiffer21adapterfusion}. Both approaches aimed to select between shared representation and task-specific representations. The proximity between tasks is not taken into account in the routing mechanism. Our method proposed a different approach to the problem of multi-task routing in the underlying network.

\section{Conclusions and outlook}\label{sec:conclusion}
In this work, we have presented a novel method for multdomain and multilingual translation. It allows us to jointly search for an optimal assignment of sub-networks to tasks and to learn the parameters of the underlying network. Our method relies on a sound mathematical framework and an end-to-end optimization procedure which implies very little extra parameters. The training course is reasonably longer (100k iterations in multidomain translation) than training standard models while considering the gain in performance.\fyDone{There is a computational cost = 100k iterations} Experimentally, we achieve a large improvement over Transformer baseline without zero additional parameters; our performance are also comparable to that of a strong a multi-task baseline using residual adapter modules which however includes a large number of extra parameters. For multilingual translation, our model outperforms multilingual Transformer and Language Adapters in 3 our of 4 cases. Besides, we provided an thorough analysis of the similarities between learned sub-networks and demonstrated a strong correlation between the sub-networks' similarity and the proximity of the corresponding tasks (domain or language).

@@

\begin{align*}
&\frac{\partial g}{\partial \bar{\nu}} \times \frac{\partial \bar{\nu}}{\partial \Phi_l^d} + \frac{\partial g}{\partial \Phi_l^d} = 0 \\
& \Rightarrow \frac{\partial \bar{\nu}}{\partial \Phi_l^d} = - \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}
\end{align*}

@@

\section{Appendix B}
\label{appendix:b}
In this section, we give a simple proof of Inequality \ref{eq:entropy}. In fact, we only need to prove $\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] \geqslant \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]$. The proof is as follows
\begin{align*}
&\mathbb{H} \big[ P(i_1,\cdots,i_k | \Phi_l^d) \big] = - \displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log}P(i_1,\cdots,i_k | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \displaystyle{\sum_{j=2}^k}\text{log}P(i_j | i_1,\cdots,j_{j-1},\Phi_l^d) +  \text{log} P(i_1 | \Phi_l^d) \big] \\
&\geqslant -\displaystyle{\mathop{\mathbb{E}}_{i_1,\cdots,i_k | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] \\
&= -\displaystyle{\mathop{\mathbb{E}}_{i_1 | \Phi_l^d}} \big[ \text{log} P(i_1 | \Phi_l^d) \big] = \mathbb{H} \big[ P(i_1 | \Phi_l^d) \big]
\end{align*}
\section{Appendix C}
\begin{breakablealgorithm}
\caption{Training \system{LaMGD}} \label{alg:mdmt}
\begin{algorithmic}[1]
\REQUIRE {
  \begin{itemize}
    \setlength{\itemsep}{-1pt}    \setlength{\parsep}{0pt}
  \item $n_d$ corpora $C^d, d \in [1, \dots,n_d]$ for $n_d$ domains equipped by an empirical distribution $D_d(x)$
  \item number of retained groups: k, number of groups: $n_p$
  \item $i=0$
 \end{itemize}}
\REPEAT 
\STATE{Pick a batch from domain $d$}
\STATE{Sampling $\forall p,  g_l^d(p) \displaystyle{\mathop{\sim}^{\text{i.i.d}}} \operatorname{Gumbel}(0,1)$}
\STATE{Solve the equation
\begin{align*}
\displaystyle{\mathop{\sum}_{i=1}^{n_p}} \sigma(\frac{g_l^d(i) + \Phi_l^d(i) + \nu}{\tau}) = k
\end{align*}
using binary search}
\STATE{Computing mask of each layer
\begin{align*}
\hat{m}_l^d(\tau) &= \sigma(\frac{g_l^d + \Phi_l^d + \bar{\nu}}{\tau})
\end{align*}}
\STATE{Applying computed masks to their corresponding layer
\begin{align*}
  \forall l \in [0,\cdots, L-1]: \tilde{h}^l &= h^l \odot r_l^d ,\\
  h^{l+1} &= \operatorname{LAYER}^{l+1}(\tilde{h}^l) ,
\end{align*}
}
\STATE{Compute gradient of training loss over the underlying Transformer
$$\Delta_{\theta} = \frac{\partial L}{\partial \theta}$$}
\STATE{Compute gradient over the $\operatorname{Soft-Top-K}$ masks
$$\frac{\partial D}{\partial \hat{m}_l^d(\tau)}$$}
\STATE{Compute implicit gradient of the $\operatorname{Soft-Top-K}$ masks over $\Phi^d$
$$- \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d}$$}
\STATE{Compute the gradient the training over $\Phi$
$$ - \frac{\partial D}{\partial \hat{m}_l^d(\tau)} \times \big(\frac{\partial g}{\partial \bar{\nu}}\big)^{-1} \times \frac{\partial g}{\partial \Phi_l^d} + \frac{\partial \mathbb{H} \big[ \operatorname{softmax}(\Phi_l^d)\big] }{\partial \Phi_l^d} $$}
\STATE{Update $\theta$ and $\Phi^d$ with their gradients}
\STATE{$i=i+1$}
\UNTIL{$i> iter\_num$}
\end{algorithmic}
\end{breakablealgorithm}
\section{Appendix D}
\begin{figure*}[h!]
\includegraphics[width=0.7\textwidth]{multi_domain}
\caption{Visualization of domains according to their dropouts (a large vector concatenating the dropping masks of all the layers of the model) constructed by PCA.}
\label{fig:domain}
\end{figure*}
\end{document}
